---
title: "01 - Minimising Mutual Information"
output:
  html_document:
    toc: true
    toc_float: true
---

In this example, I show how you can minimise the mutual information between a continuous independent variable, and a continuous confounding variable. This lets you design controlled stimuli with continuous predictors.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev.args = list(png = list(type = "cairo")))
ggplot2::theme_set(ggplot2::theme_bw())
```

# Setup

## Load Packages

```{r, message=FALSE}
library(tidyverse)
library(infotheo)  # package with functions for calculating mutual information
```

## Import the Data

```{r, message=FALSE}
stim_pool <- read_csv("stim_pool.csv")
```

# Minimising Mutual Information

Let's create a list of 100 items, where age varies continuously, but face width is controlled for.

### **Step 1**: Create a vector of seeds

Before we can start with the simulations, we need to make sure our datasets will be reproducible. We can do this by creating a vector of seeds which will be used to select our stimuli. Random seeds make the random number generator give reproducible results. If we want to run 50000 simulations, this code will create a vector of 50000 random seeds for us. Random seed values can be any integer (whole number) we can represent in R. The smallest possible seed is 1, and we can use `.Machine$integer.max` to get the maximum possible seed number R will let us represent.

I also set the seed first so the result are reproducible, but feel free to comment that out if you would like to get a different result to mine.

```{r}
set.seed(42)  # comment this line out (put # at the start of the line) to get a different result to mine
n_iter <- 50000
seeds <- sample(1:.Machine$integer.max, n_iter)
```

### **Step 2**: Simulate a large number of samples

We want to simulate a large number of random samples. On each iteration, we can record the mutual information between face width and age. We can use `map_df()` to loop over our seeds and record the mutual information values from each iteration in a large dataframe.

Depending on how many iterations we run, this could take a long time.

```{r}
res <- map_df(seeds, function(seed_i) {
  # set the seed, so we can recreate any sample later
  set.seed(seed_i)
  
  # get a random sample of 200 items from the pool
  sample <- slice_sample(stim_pool, n = 100)
  
  # discretise the face_width values, since the mutinformation methods need discrete data
  face_width_disc <- discretize(sample$face_width)
  
  # get the mutual information between age and face width
  mi_value <- mutinformation(sample$age, face_width_disc)
  
  # return a row of the dataframe with the seed and overlap value
  tibble(seed = seed_i, mi = mi_value)
})
```

### **Step 3**: Check the results

We can now look at the distribution of mutual information values from all the random samples. The closer to 0, the less mutual information in the variables, and the better the match. I've added the points below the distribution to show the results from each individual sample.

```{r}
ggplot(res, aes(mi)) +
  geom_density() +
  geom_point(aes(y = -0.5), position = position_jitter(height=0.5)) +
  xlab("Mutual Information")
```

### **Step 4**: Find the best stimulus set

We can now sort the results by mutual information (ascendingly) to order them by how well-matched the conditions are.

```{r}
res_sorted <- arrange(res, mi)

res_sorted
```

Now we can just extract the best seed as that which has the lowest mutual information.

```{r}
best_seed <- res_sorted %>%
  pull(seed) %>%
  first()

best_seed
```

### **Step 6**: Recreate the best stimulus set

Now we know that *`r best_seed`* is the best seed, we can use that seed to recreate the stimulus set. This code needs to be identical to our simulation code to ensure we recreate the exact same stimulus set.

```{r}
set.seed(best_seed)
  
best_stim <- slice_sample(stim_pool, n = 100)
```

To prove that the stimuli are well-controlled, here is a scatter pot of `age` (the IV) by `face_width` (the control variable):

```{r}
ggplot(best_stim, aes(age, face_width)) +
  geom_point() +
  geom_smooth()
```

We could use the faces in `best_stim` for our experiment, and be fairly sure that we've controlled for face width.

For comparison, here is what the worst stimulus set looks like:

```{r}
# sort descendingly (highest mutual information first), and extract the top seed
worst_seed <- res_sorted %>%
  arrange(desc(mi)) %>%
  pull(seed) %>%
  first()

set.seed(worst_seed)

# recreate the worst stimulus set
worst_stim <- slice_sample(stim_pool, n = 100)

ggplot(worst_stim, aes(age, face_width)) +
  geom_point() +
  geom_smooth()
```
