---
title: "01 - Item-Wise Matching"
output:
  html_document:
    toc: true
    toc_float: true
---

When you match item-wise, you're finding items in each condition which are directly comparable across control variables. In this example, I'll show a simple item-wise example controlling for one variable.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev.args = list(png = list(type = "cairo")))
ggplot2::theme_set(ggplot2::theme_bw())
```

# Setup

## Load Packages

```{r, message=FALSE}
library(tidyverse)
library(LexOPS)  # package with functions for item-wise matching
```

## Import the Data

```{r, message=FALSE}
stim_pool <- read_csv("stim_pool.csv")
```

# Item-Wise Matching with LexOPS

### **Step 1**: Find Matches

Let's create a list of 100 young and 100 old faces, matched on face width within ±0.5 pixels. We can do this all in one go with LexOPS.

1. Use `set_options()` to tell LexOPS which column contains the stimulus IDs
2. Use `split_by()` to specify the conditions
3. Use `control_for()` to specify the control variable
4. Use `generate()` to find 100 matches, with a seed for reproducibility

```{r}
stim <- stim_pool %>%
  set_options(id_col = "stim_id") %>%
  split_by(age, 1:28 ~ 50:100) %>%
  control_for(face_width, -0.5:0.5) %>%
  generate(n=100, seed=42)
```

This gives us a list of matched pairs of young and old faces. Each row has a young face `A1`, and old face `A2`, and `match_null` (usually only relevant for designs with >2 conditions - see [04 - Setting the Match Null](iw_04_match_nulls.html)):

```{r, eval=FALSE}
stim
```

```{r, echo=FALSE}
as_tibble(stim)
```

### **Step 2**: Check Stimuli

We can quickly check how well our stimuli are matched with the `plot_design()` function.

```{r}
plot_design(stim)
```

This shows the expected result, where:

* Condition A1 is all young faces, while condition A2 is all old faces
* Condition A1 and A2 are matched pairwise for face width - note that this also results in very similar distributions
