---
title: "03 - More than 2 Conditions"
output:
  html_document:
    toc: true
    toc_float: true
---

In this example I'll show how you can add more levels to an independent variable.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev.args = list(png = list(type = "cairo")))
ggplot2::theme_set(ggplot2::theme_bw())
```

# Setup

## Load Packages

```{r, message=FALSE}
library(tidyverse)
library(LexOPS)  # package with functions for item-wise matching
```

## Import the Data

```{r, message=FALSE}
stim_pool <- read_csv("stim_pool.csv")
```

# Matching >2 Conditions

### **Step 1**: Find Matches

This time, we use the same design as in [02 - Multiple Control Variables](iw_02_multiple_controls.html), but with an extra middle-aged condition. In total, we'll have three conditions:

* Younger faces, aged <=28
* Middle-Aged faces, aged 35-45
* Older faces, aged >=50

We can add new age conditions with the `~` symbol:

```{r}
stim <- stim_pool %>%
  set_options(id_col = "stim_id") %>%
  # create the three levels of age
  split_by(age, 1:28 ~ 35:45 ~ 50:100) %>%
  # control for all variables
  control_for(face_width, -0.5:0.5) %>%
  control_for(face_length, -2:2) %>%
  control_for(task_rt, -25:25) %>%
  control_for(gender) %>%
  generate(n=100, seed=42)
```

This gives us a list of matched triplets of young, middle-aged, and old faces. Each row has a young face `A1`, a middle-aged face `A2`, and an old face `A3`. There is also a `match_null` column (see [04 - Setting the Match Null](iw_04_match_nulls.html) for an explanation of match nulls).

```{r, eval=FALSE}
stim
```

```{r, echo=FALSE}
as_tibble(stim)
```

### **Step 2**: Check Stimuli

We can check the values for each item with `long_format()`:

```{r, eval=FALSE}
long_format(stim)
```

```{r, echo=FALSE}
as_tibble(long_format(stim))
```

We can look at the quality of the matching of the numeric values with `plot_design()`:

```{r}
plot_design(stim)
```

This shows the expected result, with the control variables matched item-wise across the three conditions.
