---
title: "08 - Item-Wise Matching from Scratch"
output:
  html_document:
    toc: true
    toc_float: true
---

In all the previous examples, I use LexOPS to match item-wise. In this example, I show code for item-wise matching without using LexOPS. The idea is to show how item-wise matching with tolerances actually works.

As a simple example, we'll recreate the design from [01 - Item-Wise Matching](iw_01_item-wise.html) using hand-written code.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev.args = list(png = list(type = "cairo")))
ggplot2::theme_set(ggplot2::theme_bw())
```

# Setup

## Load Packages

```{r, message=FALSE}
library(tidyverse)
```

## Import the Data

```{r, message=FALSE}
stim_pool <- read_csv("stim_pool.csv")
```

# Item-Wise Matching from Scratch

Let's create a list of 100 young and 100 old faces, matched on face width.

### **Step 1**: Identify conditions

First, we define our conditions. Let's say young faces are <28 years old, and old faces are >50 years old. We can use dplyr's `mutate()` and `case_when()` to create our conditions.

```{r}
stim_pool <- mutate(stim_pool, cond = case_when(age<28 ~ "young", age>50 ~ "old"))
```

This gives us >1000 candidate items for each of our conditions.

```{r}
count(stim_pool, cond)
```

### **Step 2**: Decide on Tolerance

The first step is to decide how *closely* we want to match face width. Let's imagine that we decide to match face width within ±0.5 pixels. For reference later, we'll store this in a vector:

```{r}
tol <- c(-0.5, 0.5)
```

### **Step 3**: Find Matches

Here is the code for finding 100 pairs of young and old faces matched for face width.

On each iteration, we:

1. Get a list of all candidate faces which haven't already been selected.
2. Pick a random item from the young condition.
3. Try pick a random item from the old condition which is suitably close (within ±0.5 face width).
4. If there was no suitable match, skip to the next iteration. Otherwise, record the match.

We keep going until we have `n_stim` (100) matched pairs. To do this, we can use a `while` loop. Normally you would also want to add `break` statements to handle cases when the method fails, but I've left these out for now to keep the code simple.

This will give us a list of matched pairs in the form `list(c("stim_42", "stim_54"), c("stim_972", "stim_43"))`, of length 100.

```{r}
# number of stimuli we want
n_stim <- 100

# the conditions we're interested in
age_conds <- c("young", "old")

# keep track of how many matches have been found
n_found <- 0
matches <- list()

while (n_found < n_stim) {
  # 1. filter the pool to exclude items already matched
  iter_pool <- filter(stim_pool, !stim_id %in% unlist(matches))
  
  # 2. pick a cond_i item at random, and get it's data
  young_item <- iter_pool %>%
    filter(cond == "young") %>%
    slice_sample(n = 1)
  
  # 3. filter the pool to find an item suitably close to the match null
  old_item <- iter_pool %>%
    filter(
      cond == "old",
      face_width >= young_item$face_width + tol[1],
      face_width <= young_item$face_width + tol[2]
    ) %>%
    slice_sample(n = 1)
  
  # 4. if no suitable match, end the iteration...
  if (nrow(old_item) == 0) next
  
  # ...otherwise, record the match
  matches[[length(matches) + 1]] <- c(young = young_item$stim_id, old = old_item$stim_id)
  n_found <- n_found + 1
}
```

### **Step 4**: Check Stimuli

Now we have a list of all the matched pairs, we can put this in a dataframe like so .

```{r}
match_df <- map_df(1:length(matches), function(i) {
  tibble(
    pair_nr = i,
    young = matches[[i]]["young"],
    old = matches[[i]]["old"]
  )
})

match_df
```

We can add to this the face width values associated with each young and old item, and calculate the difference.

```{r}
stim_pool_young <- stim_pool %>%
  rename(face_width_young = face_width) %>%
  select(stim_id, face_width_young)

stim_pool_old <- stim_pool %>%
  rename(face_width_old = face_width) %>%
  select(stim_id, face_width_old)

match_df <- match_df %>%
  left_join(stim_pool_young, by=c("young" = "stim_id"))  %>%
  left_join(stim_pool_old, by=c("old" = "stim_id")) %>%
  mutate(face_width_diff = face_width_young - face_width_old)

match_df
```

Here is the distribution of differences between the young and old faces. Notice that they are between `r tol[1]` and `r tol[2]`.

```{r}
match_df %>%
  ggplot(aes(face_width_diff)) +
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(n.breaks=10)
```

We can also create a plot similar to `plot_design()` to show the distributional similarity at the same time:

```{r}
match_df %>%
  select(young, old, pair_nr) %>%
  pivot_longer(c(young, old), names_to="cond", values_to="stim_id") %>%
  left_join(select(stim_pool, stim_id, face_width), by="stim_id") %>%
  ggplot(aes(cond, face_width)) +
  geom_violin() +
  geom_point() +
  geom_line(aes(group = pair_nr))
```

# Summary

I decided to write LexOPS because it lets you match items with readable and easily editable code, but as you can see the method is actually very simple! You could adapt the code written here to create any stimuli you could create with LexOPS, including the other examples provided.
