---
title: "05 - Factorial Designs"
output:
  html_document:
    toc: true
    toc_float: true
---

In a factorial design, rather than splitting on a single variable, it is common to manipulate multiple independent variables to create a design that allows you to test an interaction. In this example, I show how you might create stimuli for a 3 (age: young, middle, old) x 2 (gender: female, male) design.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev.args = list(png = list(type = "cairo")))
ggplot2::theme_set(ggplot2::theme_bw())
```

# Setup

## Load Packages

```{r, message=FALSE}
library(tidyverse)
library(LexOPS)  # package with functions for item-wise matching
```

## Import the Data

```{r, message=FALSE}
stim_pool <- read_csv("stim_pool.csv")
```

# Generating Stimuli for Factorial Designs

### **Step 1**: Find Matches

Here, we'll create stimuli for a 3 (age: young, middle, old) x 2 (gender: female, male) design, controlling for face width and face length. To split by both independent variables, we can just call `split_by()` an additional time:

```{r, warning=FALSE}
stim <- stim_pool %>%
  set_options(id_col = "stim_id") %>%
  # create the three levels of age
  split_by(age, 1:28 ~ 35:45 ~ 50:100) %>%
  split_by(gender, "f" ~ "m") %>%
  # control for all variables
  control_for(face_width, -1:1) %>%
  control_for(face_length, -1:1) %>%
  generate(n=100, seed=42, match_null="inclusive")
```

This gives us a list of matched sets of faces for six conditions:

* Young female faces (A1_B1)
* Young male faces (A1_B2)
* Middle-aged female faces (A2_B1)
* Middle-aged male faces (A2_B2)
* Old female faces (A3_B1)
* Old male faces (A3_B2)

```{r, eval=FALSE}
stim
```

```{r, echo=FALSE}
as_tibble(stim)
```

### **Step 2**: Check Stimuli

We can check the values for each item with `long_format()`:

```{r, eval=FALSE}
long_format(stim)
```

```{r, echo=FALSE}
as_tibble(long_format(stim))
```

We can look at the quality of the matching of the numeric values with `plot_design()`:

```{r}
plot_design(stim)
```

This shows the expected result, with all six conditions matched for face width and face length, whereas age and gender differ.
