---
title: "02 - Multiple Controls"
output:
  html_document:
    toc: true
    toc_float: true
---

In this example I'll show how you can match multiple control variables, some numeric and some categorical.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev.args = list(png = list(type = "cairo")))
ggplot2::theme_set(ggplot2::theme_bw())
```

# Setup

## Load Packages

```{r, message=FALSE}
library(tidyverse)
library(LexOPS)  # package with functions for item-wise matching
```

## Import the Data

```{r, message=FALSE}
stim_pool <- read_csv("stim_pool.csv")
```

# Using Multiple Controls

### **Step 1**: Find Matches

To control for multiple variables in one stimulus set, we can just add more calls to `control_for()`. We'll control for:

* Face width, within ±0.5 pixels
* Face length, within ±2 pixels
* Average response time on a behavioural task, within ±25 ms
* Gender (match faces with a face of the same gender)

```{r}
stim <- stim_pool %>%
  set_options(id_col = "stim_id") %>%
  split_by(age, 1:28 ~ 50:100) %>%
  control_for(face_width, -0.5:0.5) %>%
  control_for(face_length, -2:2) %>%
  control_for(task_rt, -25:25) %>%
  control_for(gender) %>%
  generate(n=100, seed=42)
```

This gives us a list of matched pairs of young and old faces. Each row has a young face `A1`, and old face `A2`, and `match_null` (usually only relevant for designs with >2 conditions - see [04 - Setting the Match Null](iw_04_match_nulls.html)):

```{r, eval=FALSE}
stim
```

```{r, echo=FALSE}
as_tibble(stim)
```

### **Step 2**: Check Stimuli

We can check the values for each item with `long_format()`:

```{r, eval=FALSE}
long_format(stim)
```

```{r, echo=FALSE}
as_tibble(long_format(stim))
```

We can look at the quality of the matching of the numeric values with `plot_design()`:

```{r}
plot_design(stim)
```

This shows the expected result, where:

* Face width is matched closely, within ±0.5 pixels
* Face length is matched less closely, within ±2 pixels
* Average response time on a behavioural task is matched very closely, within ±25 ms
